---
title: "An Animated Attrition Analysis"
author: "Andrew Iraola"
date: "March 5, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Loading Necessary Packages    
```{r message=FALSE}
library(tidyverse)
library(GGally)
library(readxl)
library(broom)
library(stargazer)
library(modelr)
library(gganimate)
library(gifski)
library(ggthemes)
```


### Pulling the Necessary Data from Analytix
Let's wrangle the (excel spreadsheet) datasets into R Markdown. For this project, we're going to utilize Attrition and Enrollment Data from 2011-2018.
The only change I made to the spreadsheet(s) was to delete row 2, so that only the variable names considered and not their descriptors. 

```{r}
attrition <- read_excel("C:/Users/airaola/Desktop/TidyTuesdays/Week 1/DataSet Attrition .xlsx")
JD1_enrollment <-read_excel("C:/Users/airaola/Desktop/TidyTuesdays/Week 1/DataSet Enrollment .xlsx", sheet=2)
Total_JD_enrollment <-read_excel("C:/Users/airaola/Desktop/TidyTuesdays/Week 1/DataSet Enrollment .xlsx")
```


I'm now going to subset the second worksheet of the `Dataset Enrollment` dataset to focus on only the variables (column names) we need. 
```{r}
JD1_enrollment <- JD1_enrollment %>%
  dplyr::select(schoolid, calendaryear, totaljd1)
```


Repeating the process for the first worksheet (*Total JD Enrollment*)
```{r}
Total_JD_enrollment <- Total_JD_enrollment %>%
  dplyr::select(schoolid, calendaryear, totaljd)
```


Before we add these to the Attrition Dataset, "Dataset Attrition", we need to tidy up a few things.
Firstly, it seems to be reading in `schoolid` as a character variable. Let's switch that to numeric with the below command:
```{r}
attrition$schoolid <- as.numeric(attrition$schoolid)
```

Let's also add a column to the Attrition Dataset that calculates "Total Attrition" as follows:
```{r}
attrition %>%
mutate(TotalAttrition = totaljd1attrition + totaljd2attrition + totaljd3attrition + totaljd4attrition, na.rm=T)

```


Now, let's merge the three datasets! To do this, I'm going to add the columns we pulled from `JD1_enrollment` and `Total_JD_enrollment` to the `attrition` dataset we just got done manipulating. The merge will be by `schoolid` and `calendaryear` so that everything stays in it's corresponding row (school) and year according. 
```{r}
attrition <- attrition %>%
  left_join(JD1_enrollment, by = c("schoolid", "calendaryear")) %>%
  left_join(Total_JD_enrollment, by = c("schoolid", "calendaryear"))
```



And now, the fun part- We'll begin with grouping by `calendaryear`, then summing each column for which we want yearly totals. 

1. JD1 Attrition 
2. Total Attrition
3. JD1 Enrollment
4. Total Enrollment

We'll then `mutate` these sums to get the proportion of students that left law school based on enrollment figures. We'll do this for JD1 as well as the total (1L-4L).

Filtering out 2018 (for which there is not yet data), we then plot the two mutations: `JD1_AttritionPercentage` and `Total_AttritionPercentage`.
I chose to use a `theme_minimal()` line graph for now.
```{r}
attrition %>%
  group_by(calendaryear) %>%
  summarise(JD1_attrition = sum(totaljd1attrition, na.rm=T),
            JD2_attrition = sum(totaljd2attrition, na.rm=T),
            JD3_attrition = sum(totaljd3attrition, na.rm=T),
            JD4_attrition = sum(totaljd4attrition, na.rm=T),
            JD1_enrollment = sum(totaljd1, na.rm=T),
            Total_JD_enrollment = sum(totaljd, na.rm=T)) %>%
  mutate (Total_AttritionPercentage = ((JD1_attrition + JD2_attrition + JD3_attrition + JD4_attrition)/Total_JD_enrollment)*100) %>%
  mutate (JD1_AttritionPercentage = (JD1_attrition/JD1_enrollment)*100) %>%
  filter (calendaryear != 2018) %>%
  
  ggplot (aes(calendaryear, JD1_AttritionPercentage)) +
  geom_line (aes(calendaryear, JD1_AttritionPercentage), color="forest green") +
  geom_line (aes(calendaryear, Total_AttritionPercentage, color="red")) + 
  scale_color_discrete(name = "Type", labels = c("Total", "First")) +
  theme_minimal() +
  xlab("Year") +
  ylab("Percent") +
  ylim(0,15) + 
  ggtitle("Attrition as a Percentage of Enrollment", "March 2018") + 
  theme(plot.title=element_text(hjust = 0.5)) + 
  theme(plot.subtitle = element_text(hjust = 0.5)) 
    
```

But that doesn't look quite right. It seems to only register the secondary label-heading and ignore the first. 
Let's try something different with the `select` and `gather` functions. I'll also change from the minimal theme to that of the Economist. 

```{r}
P1 <- attrition %>%
  group_by(calendaryear) %>%
  summarise(JD1_attrition = sum(totaljd1attrition, na.rm=T),
            JD2_attrition = sum(totaljd2attrition, na.rm=T),
            JD3_attrition = sum(totaljd3attrition, na.rm=T),
            JD4_attrition = sum(totaljd4attrition, na.rm=T),
            JD1_enrollment = sum(totaljd1, na.rm=T),
            Total_JD_enrollment = sum(totaljd, na.rm=T)) %>%
  mutate (Total_AttritionPercentage = ((JD1_attrition + JD2_attrition + JD3_attrition + JD4_attrition)/Total_JD_enrollment)*100) %>%
  mutate (JD1_AttritionPercentage = (JD1_attrition/JD1_enrollment)*100) %>%
  filter (calendaryear != 2018) %>%
  
  select(calendaryear, Total_AttritionPercentage, JD1_AttritionPercentage) %>%
  gather(key, value, -calendaryear) %>%
  ggplot(aes(calendaryear, value, color=key, frame = calendaryear)) +
  geom_line() + 
  scale_color_discrete(name = "Type", labels = c("Total", "First Year")) +
  theme_economist() +
  xlab("Year") +
  ylab("Percent") +
  ylim(0,15) + 
  ggtitle("Total Law School Attrition as a Percentage of Enrollment", "March 2018") + 
  theme(plot.title=element_text(hjust = 0.5)) + 
  theme(plot.subtitle = element_text(hjust = 0.5)) 

P2 <- P1 + geom_point(aes(group=seq_along(calendaryear))) +
  transition_reveal(calendaryear)
P2
```

Now let's get it ready for presentation.
**`gganimate`**, do yo thing!


```{r}
P2 <- P1 + geom_point(aes(group=seq_along(calendaryear))) +
  transition_reveal(calendaryear)
P2
```

```{r}
anim_save("TidyTuesday1")
```