---
title: "Tidy Tuesday Week Two"
author: "Andrew Iraola"
date: "March 19, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Let's Get at that Dataset

```{r echo = FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(Amelia)
library(GGally)
library(broom)
library(modelr)
library(stargazer)
library(car)
library(lme4)
library(htmlwidgets)
library(ggplot2)
library(plotly)
library(lattice)
```

```{r echo=FALSE, message=FALSE}
schools <- lapply(Sys.glob("R:/Level 2/17-18_School*.xlsx"), read_excel)

all_schools <- Reduce(function(x, y) merge(x, y, all=TRUE), schools, accumulate=FALSE)
```

Checking for missingness...
```{r}
missmap(all_schools)
```

...Yea that didn't tell me anything, 
it kinda looks like a [Rothko](https://www.google.com/search?q=rothko&rlz=1C1GCEA_enUS793US793&source=lnms&tbm=isch&sa=X&ved=0ahUKEwivob770Y7hAhWjiOAKHUzAA1MQ_AUIDigB&biw=1497&bih=658)

```{r message=FALSE, echo=FALSE}

# Race as a factor
all_schools$Race <- ifelse (all_schools$Race=="Black or African American" | all_schools$Race=="Black or African                               american", paste("Black"), all_schools$Race)
all_schools$Race <- ifelse (all_schools$Race=="Hispanics of any race", paste("Hispanic"), all_schools$Race)
all_schools$Race <- ifelse (all_schools$Race=="White", paste("White"), all_schools$Race)
all_schools$Race <- ifelse (all_schools$Race=="Asian", paste("Asian"), all_schools$Race)
all_schools$Race <- ifelse (all_schools$Race=="American Indian or Alaska Native"|all_schools$Race=="Native                                    Hawaiian or Other Pacific Islander", paste("NA/PI"), all_schools$Race)
all_schools$Race <- ifelse (all_schools$Race=="Nonresident alien"| all_schools$Race=="Two or More                                             Races"|all_schools$Race=="Race and Ethnicity Unknown", paste("Other"), all_schools$Race)
all_schools$Race <- factor(all_schools$Race)
all_schools$Race <- factor(all_schools$Race, levels=c("White", "Asian", "Black", "Hispanic", "NA/PI", "Other"))
```


Let's try something different. Instead, only focusing on all our relevant variables for analysis. 
```{r}
all_schools %>%
  select(BarResult, Race, Gender, CreditHrs_Doctrinal, CreditHrs_Skill, CreditHrs_Clinic, CreditHrs_Externship,
         CreditHrs_Journals, CreditHrs_Moot, CreditHrs_Trial, CreditHrs_Other) %>%
missmap()
```
Now, although we can't see any specifity on the Y-axis, we can still contextualize how much missingness there is for our variable-set. Given that only 2 percent is missing, that tells us we can continue with our analysis.


Now, let's peak at the relationship between `Bar Result` and various classes for which credit hours were reported.

```{r}
sum(is.na(all_schools$BarResult)) 
```

In isolation:
```{r}
all_schools %>%
  ggplot(aes(x = CreditHrs_Skill, y = BarResult, color = as.factor(SchoolID))) +
  geom_point()
  #geom_smooth(method = "lm")
```

SchoolID as factor
```{r}
all_schools$SchoolID <- as.factor(all_schools$SchoolID)
```



## Multi-level Model

```{r}
all_schools <- all_schools %>%
  mutate(BarPass = ifelse(BarResult == "Pass", 1, 0)) 

BarResult_Model <- glm (BarPass ~ CreditHrs_Doctrinal + CreditHrs_Skill + Gender + Race, 
                             data = all_schools)
summary(BarResult_Model)
```
Well that's wholly uninteresting...


```{r}
all_schools 
  
  p <- ggplot (all_schools, aes(x = Yr1_Rank, y = Final_GPA, color = as.factor(SchoolID))) +
  geom_point () +
  xlab ("1L Class Ranking") +
  ylab ("Final GPA") +
  ggtitle ("Final Law School GPA by 1L Class Ranking", "March 2019") +
  scale_color_discrete(name = "School") + 
  geom_smooth (method = "lm")

p <- ggplotly(p)

saveWidget(p,"TidyTuesday2.html")
```

```{r}
all_schools <- all_schools %>%
  mutate(BarPass = ifelse(BarResult == "Pass", 1, 0)) 

BarResult_Model <- lm(BarPass ~ clquest + clunprep + journal + mootcrt + orgmembr + 
                          work + caredep + commute + parenteduc + grades + Race + Gender, 
                             data = all_schools)
summary(BarResult_Model)
```










