---
title: 'Tidy Tuesday #4: Admissions'
author: "Andrew Iraola"
date: "April 24, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Alright, Let's Get Into It..

```{r message=FALSE}
### Loading Necessary Packages    
library(tidyverse)
library(broom)
library(Amelia)
library(GGally)
library(readxl)
library(car)
library(stargazer)
library(ggplot2)
library(modelr)
library(caret)
library(magick)
library(gganimate)
library(stargazer)
```

## Loading Our Data in

```{r}
Admissions <- read_excel("Admissions Data.xlsx")
```

```{r}
summary(Admissions)
```

To `tidy` up race are recoding it to become a factor that is much easier to read for both the analyst and general audience. 
In this case, we have only 6 non-white students in total (13.6% of the sample) so we are going to separate race by the two largest categories, `White` and `Non-White`. 

```{r}
Admissions$Race <- ifelse(Admissions$Ethnicity_1_Long=="Black/African American", paste("African American"), Admissions$Ethnicity_1_Long)


Admissions$Race <- ifelse(Admissions$Ethnicity_1_Long=="Hispanic/Latino"| Admissions$Ethnicity_1_Long=="Puerto Rican", paste("Hispanic"), Admissions$Race)

Admissions$Race <- ifelse(Admissions$Ethnicity_1_Long=="Caucasian/White", paste("White"), Admissions$Race)

Admissions$Race <- ifelse(Admissions$Ethnicity_1_Long=="American Indian/Alaskan Native"|Admissions$Ethnicity_1_Long=="Native Hawaiian/Other Pacific Islander", paste("Other"), Admissions$Race)


Admissions$Race <- ifelse(Admissions$Ethnicity_1_Long=="Other"| Admissions$Ethnicity_1_Long=="Did not Indicate", paste("Other"), Admissions$Race)

 Admissions$Race <- factor(Admissions$Race)
 
 Admissions$Race <- factor(Admissions$Race, levels=c("White", "African American", "Hispanic", "Asian", "Other"))
 
table(Admissions$Race)

```
Before running this model, let's convert the BarResult` variable into a number variable.

```{r}
Admissions <- Admissions %>%
  mutate(Admission_Status = ifelse(Decision_Status_Long == "Admitted", 1, 0)) %>%
  mutate(Law_Based_Major = ifelse(Undergraduate_Major_Long=="LAW"| Undergraduate_Major_Long=="PRE-LAW", 1, 0)) %>%
  mutate(Male = ifelse(Gender_Applicant == "M", 1, 0))
```


# Outcome 1: Race, Gender, LSAT and other variables on Admission Status

DV:

1. Admission Status (`Admission_Status`)

IVs:

1. Race/ethnicity (`Race`)
2. Gender (`Male`)
3. LSAT (`High_LSAT_Score`)
4. UGPA (`Cumulative_GPA`)
5. Undergraduate Major (`Law_Based_Major`)

Running the first model:

```{r}
Admissions_Model <- glm(Admission_Status ~ Race + Male + High_LSAT_Score + Cumulative_GPA + Law_Based_Major, family="binomial", data=Admissions)
summary(Admissions_Model)
```


```{r}
vif(Admissions_Model)
```




## Full Analysis Predictions


### LSAT Score

After we create the data frame, well will create the predictions as a new variable within the dataframe. 

```{r}
LSAT_admit <- Admissions %>%
  data_grid(High_LSAT_Score = seq_range(High_LSAT_Score, 100), Race, .model = Admissions_Model)

LSAT_admit <- LSAT_admit %>%
  mutate(pred = predict(Admissions_Model, LSAT_admit, type = "response"))
```

```{r}
P <- ggplot(LSAT_admit, 
       aes(x = High_LSAT_Score, y = pred,
           color = Race)) +
  geom_point() +
  xlab("LSAT Score") +
  ylab("%") +
  labs(title = "Predicted Probabilities of Being Admitted to Law School (2013-17)", subtitle = "Model accounts for Demographics, LSAT, UGPA and Undergraduate Major") +
  scale_colour_discrete(name = "Race/Ethnicity") +
  theme_minimal() + 
  scale_y_continuous(labels = scales::percent_format(accuracy=1)) + 
  coord_cartesian(ylim = c(0,1))  

P2 <- P + geom_point(aes(group=seq_along(High_LSAT_Score))) +
  transition_reveal(High_LSAT_Score)  

gif1 <- animate(P2, fps = 10, renderer = gifski_renderer(loop = FALSE))
gif1

anim_save("TidyTuesday4")
```

## Cumulative Undergraduate GPA

After we create the data frame, well will create the predictions as a new variable within the dataframe. 

```{r}
UGPA_admit <- Admissions %>%
  data_grid(Cumulative_GPA = seq_range(Cumulative_GPA, 100), Race, .model = Admissions_Model)

UGPA_admit <- UGPA_admit %>%
  mutate(pred = predict(Admissions_Model, UGPA_admit, type = "response"))
```

```{r}
p <- ggplot(UGPA_admit, 
       aes(x = Cumulative_GPA, y = pred,
           color = Race)) +
  geom_point() +
  xlab("Cumulative GPA") +
  ylab("%") +
  labs(title = "Predicted Probabilities of Being Admitted to Law School (2013-17)", subtitle = "Model accounts for Demographics, LSAT, UGPA and Undergraduate Major") +
  scale_colour_discrete(name = "Race/Ethnicity") +
  theme_minimal() + 
  scale_y_continuous(labels = scales::percent_format(accuracy=1)) + 
  coord_cartesian(ylim = c(0,1))  

p2 <- p + geom_point(aes(group=seq_along(Cumulative_GPA))) +
  transition_reveal(Cumulative_GPA)  

gif2 <- animate(p2, fps = 15, renderer = gifski_renderer(loop = FALSE))

gif2
anim_save("TidyTuesday4.1")



```

